---
title: "Activity: Data Quality — Cleaning the Australian Salary Data"
author: "Chris"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)

# if you haven't installed lubridate, please install running the code, install.packages("lubridate")
library(lubridate)
```

# The task

The goal is to clean the dataset so that we can obtain a valid estimate
of the average salary of people in Australia.

**Data collected from users on 25/08/2026.**

### Read the raw data

```{r raw-data}
raw <- read.csv("salary_data_raw.csv")
raw
```

<br>

## Data quality issues

### Clean inconsistent and incorrect values

You may use `recode`syntax to clean the values. Once you clean, you need
to convert the datatype for some variables, using `as.numeric()` or
`dmy()` of lubridate library.

```{r}
clean <- raw %>%
  mutate(
    Salary = recode(Salary,
                    "70,000" = "70000",   # thousands separator
                    "72,000" = "72000",
                    "65k"    = "65000"),  
    Salary = as.numeric(Salary),
    Date.of.employment = recode(Date.of.employment,
      "Feb 1, 2026"     = "1/02/2026",     # written month -> dd/mm/yyyy
      "2 days ago"      = "23/08/2026",    # relative to collection date 25/08/2026
      "1/02/2000 10:00" = "1/02/2000"), # drop the time, keep the date
    Date.of.employment = dmy(Date.of.employment) # convert to datetime datatype# convert to numeric
  )

clean
```

<br>

### Dealing with the missing values

#### Approach 1 — Drop the missing values

```{r drop}
clean_drop <- clean %>%
  drop_na(Salary)

clean_drop
```

*When it is reasonable:* when the data is **missing completely at
random** and the sample is large enough that losing rows costs little.

*Cost here:* with n = 6 we throw away 17% of the sample, and we lose
Mike's age and city information as well.

<br>

#### Approach 2 — Replace with the mean

Use `ifelse`syntax to replace the missing value with mean of Salary.

```{r}
clean_mean <- clean %>%
  mutate(Salary = ifelse(is.na(Salary), mean(Salary, na.rm = TRUE), Salary))

clean_mean
```

*Cost:* it shrinks the variance and weakens any correlation with other
variables. It also ignores everything we know about Mike (55, Sydney).

<br>

#### Approach 3 — Rule-based imputation

Calculate the median salary for each city, then replace the missing
Salary value with the mean salary for that city. To get the mean salary
for each city, please use `group_by()`.

```{r rule-impute}
clean %>%
  group_by(City) %>%
  summarise(city_median = mean(Salary, na.rm = TRUE))


clean_rule <- clean %>%
  mutate(Salary = ifelse(is.na(Salary), 64750, Salary))

clean_rule
```

*Advantage:* considers the impact of other variables such as City when
imputing missing value of salary. *Cost*: the rule is the analyst's
judgement rather than something estimated from the data, and it is
coarse — it ignores Age

<br>

#### Approach 4 — Linear regression imputation

```{r}
# Run a linear regression of salary on age (lm() drops the NA row automatically)
sal_lm <- lm(Salary ~ Age + City, data = clean)

# Impute the missing salary from the fitted model
clean_lm <- clean %>%
  mutate(Salary = ifelse(is.na(Salary),
                         predict(sal_lm, newdata = clean),
                         Salary))
clean_lm
```

*Advantage:* identify the complex pattern between several predictors and
the imputed value is tailored to Mike specifically (55, Sydney).

*Cost:* overfitting

<br>

<br>

### Remaining issues

Cleaning fixed the *format* of the data. It did not fix these:

**Still unresolved in the data**

1.  **Tom's record is internally impossible** — age 27 with a start date
    in 2000 means he started work at age 1. Either the age or the date
    is wrong. Chris (started at \~15) is implausible but not impossible.
2.  **The sample cannot answer the question.** n = 6 people from two
    cities cannot estimate the average salary "of people in Australia".
    No regional, remote, industry, gender or age-group coverage; the ABS
    uses samples in the tens of thousands.
3.  **Salary is undefined.** Gross or net? Full-time or pro-rata? Does
    it include superannuation, bonuses, overtime? A part-time salary and
    a full-time salary are not comparable, and the survey does not
    record hours.
4.  **No time reference for the salary.** Is it the current salary or
    the salary at the date of employment? Tom started in 2000 — a 2000
    salary and a 2026 salary are not comparable without inflation
    adjustment.

<br>

All of the material is copyrighted under the [Creative Commons BY-SA
4.0](https://creativecommons.org/licenses/by/4.0/) copyright.
